\chapter{CEAS Error Derivation}\label{ch:ceas_error_der}

For a random variable $X$, the standard error is defined by the square root
of the second moment (also known as the standard deviation) divided by the
number of elements taken of $X$, represented by $x$.

\begin{align*}
  \Delta_x = \frac{\sigma_x}{\sqrt{n}}
\end{align*}

We estimate the standard error in a function $f$ of $X$ by taking the
partial differential of function with respect to $X$. This is known as the
\emph{delta method}.

\begin{align*}
  \Delta_f^2 =  \left( \frac{\partial f(x)}{\partial x} \Delta_x\right)^2
\end{align*}

This is a measure of the standard error of the mean value, which is
calculated by plugging the mean value $\overline{x}$ into $f$.

For a function $g$ of multiple independent variables $X_1,X_2,...,X_n$ we
can estimate the standard error $\Delta_g$ in the following way.

\marginpar{\\\bigskip if the variables are not independent we should add a
covariance term.}

\begin{align*}
  \Delta_g^2 = \sum_i^n\left(\frac{\partial f}{\partial x_i}\Delta x_i\right)^2
\end{align*}

In calculating the absorption in a CEAS experiment it is common to use the
function

\begin{align}
  \alpha(\lambda) = \left(\frac{I_0(\lambda)}{I(\lambda)}-1\right)\left(\frac{1-R(\lambda)}{d}\right)\label{eq:ceas}
\end{align}

where $I$ and $I_0$ represent the intensity of the light through the cavity
with and without the absorber present, respectively.

To calculate the squared error $\Delta_\alpha^2$, we shall use the delta
method as defined above.

\marginpar{$\lambda$ is omitted in the CEAS equation from now on for
convenience}

\begin{align*}
  \Delta_\alpha^2 &= \left(\frac{\partial \alpha}{\partial I_0}\Delta_{I_0}\right)^2  + \left(\frac{\partial \alpha}{\partial I}\Delta_I\right)^2  \\\\
                  &= \left[\frac{\Delta_{I_0}}{I}\left(\frac{1-R}{d}\right)\right]^2 + \left[\frac{I_0 \Delta_I }{I^2}\left(\frac{1-R}{d}\right)\right]^2 \\\\
                  &= \left(\frac{\Delta_{I_0}}{I} + \frac{I_0\Delta_I}{I^2}\right)^2 \left(\frac{1-R}{d}\right)^2 \\\\
           &= \left[\left(\frac{\Delta_{I_0}}{I} + \frac{I_0\Delta_I}{I^2}\right) \left(\frac{1-R}{d}\right)\right]^2
\end{align*}

This can be simplified to get the standard error.

\begin{align}
  \Delta_\alpha = \left(\frac{\Delta_{I_0}}{I} + \frac{I_0\Delta_I}{I^2}\right) \left(\frac{1-R}{d}\right)\label{eq:err}
\end{align}

This CEAS equation assumes that the reflectivity is close to 1 and that
$\alpha$ is nearly zero (only weak, low concentration absorbers). A version
of this equation that does not make this assumption is shown below.

\begin{align}
  \alpha = \frac{1}{d}\ln\left(\frac{1}{2R^2}\left(\sqrt{4R^2+\left(\frac{I_0}{I}(1-R^2)\right)^2} - \frac{I_0}{I}(1-R^2)\right)\right)\label{eq:ceas_full}
\end{align}

Here we will introduce, for convenience,

\begin{align*}
  \beta &= \sqrt{4R^2+\left(\frac{I_0}{I}(1-R^2)\right)^2}
\end{align*}

Now, we can easily find the partials

\begin{align*}
  \frac{\partial \alpha}{\partial I_0} &= \left(\dfrac{1}{I}\right)\left(\dfrac{1-R^2}{d\beta}\right)\\\\
    \frac{\partial \alpha}{\partial I} &= -\left(\dfrac{I_0}{I^2}\right)\left(\dfrac{1-R^2}{d\beta}\right)
\end{align*}

After which we can easily estimate the standard error in $\alpha$.

\begin{align*}
  \Delta_\alpha^2 &= \left(\frac{\partial \alpha}{\partial I_0} \Delta_{I_0}\right)^2  + \left(\frac{\partial \alpha}{\partial I} \Delta_{I}\right)^2 \\\\
                  &= \left(\frac{\Delta_{I_0}}{I} + \frac{I_0\Delta_{I}}{I^2}\right)^2 \left(\dfrac{1-R^2}{d\beta}\right)^2 \\\\
                  &= \left[\left(\frac{\Delta_{I_0}}{I} + \frac{I_0\Delta_{I}}{I^2}\right)\left(\dfrac{1-R^2}{d\beta}\right)\right]^2
\end{align*}

\marginpar{\\\vspace{-3em}Be careful when $I$ is zero!}

Finally, this can be simplified to the standard error

\marginpar{\\\bigskip The standard deviation version of this equation is the same except all the standard error values are replaced with their standard deviation equivalents.}

\begin{align}
  \Delta_\alpha &= \left(\frac{\Delta_{I_0}}{I} + \frac{I_0\Delta_{I}}{I^2}\right) \left(\dfrac{1-R^2}{d\beta}\right) \notag\\\notag\\
                &= \left(\frac{\Delta_{I_0}}{I} + \frac{I_0\Delta_{I}}{I^2}\right) \left(\dfrac{1-R^2}{d}\right)\left(4R^2+\left(\frac{I_0}{I}(1-R^2)\right)^2\right)^{-1/2} \label{eq:err_full}
\end{align}

Equations \eqref{eq:ceas_full} and \eqref{eq:err_full} are preferred to
\eqref{eq:ceas} and \eqref{eq:err} because neither \eqref{eq:ceas_full} nor
\eqref{eq:err_full} make assumptions on the limits of $R$ or $\alpha$,
whereas \eqref{eq:ceas} and \eqref{eq:err} do make the assumptions $R \to
1$ and $\alpha \to 0$.



\section{Notes on calculations}\label{sec:ceas_eq_notes}

While the equations for \ac{BBCEAS} and the standard deviation of the signals
is relatively straight forward, there are bounds on the calculation of
$\alpha$ that are implied.

An experimenter should ignore $\alpha = 0$, as this indicates that the
intensity measurements were indiinguishable. At such a value, a more meaningful
result would be to state that the absorption is insignificant with reference to
the LOD$_{\alpha}$.

Another problem is when the measured intensity $I$ is zero. In all of the
equations stated thus far result in a division by zero when the intensity
value is zero. The optimistic reader may state that this would never be
the case, as there will always be background counts on the detector due to
\ac{CCD} noise. However $I$ can be zero if one subtracts the background
\ac{CCD} noise from the acquired images. This subtraction is required for an
accurate measurement of $\alpha$.

This phenomenon will occur at high absorptivity values, and due to the nature
of pressure and collision broadening means that there is a high probability
that several points in a row in the intensity spectrum will be zero or nearly
zero. The resulting $\alpha$ spectrum would have a peak where the top was cut
off. While this scenario is not ideal, there is a way acquire an approximate
value for the $\alpha$ values that have been cut.

The line shapes of an electronic transition (assuming that there are no other
significant absorptions near a certain transition), after the convolution of
collision broadening, pressure broadening, and the broadening associated with
the spectrometer, take on a voigt profile. Since these broadening effects
impact all transitions equally, we can use another peak within the absorption
spectrum to calculate the parameters for this profile. It is then trivial to
scale the voigt profile to the peak that has been cut off at the top and
arrive at an approximate value of $\alpha$. The only potential problem with
this method is that all of the existing methods shown for calculating the
error in the absorption coefficient are inapplicable.
